---
title: "Final Exam"
format: 
  html:
    self-contained: true
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---

**Kennedy Lighty**

```{r}

library(tidyverse)
library(wbstats)
library(broom)
library(modelsummary)
library(marginaleffects)
library(plotly)
library(ggdist)
library(stargazer)
library(dplyr)
library(readr)
```

Please read all of the questions and tables carefully and follow all instructions. Each question has an allotted point value. Be as thorough and specific as possible; extra calculations and incorrect information, even in the presence of correct information, will result in point deductions. Be sure to show all formulas and all necessary work to answer the questions. You may upload your completed exam to the Elms course site.

::: callout-note
While this is an open-note exam, you are not to receive assistance from anyone else. As usual, the Honor Code applies.
:::

**Total points: 50**

## Conceptual Questions

Please include all work (and computations) necessary to answer the questions.

**Total points: 20 (2 points each)**

### Question 1

Suppose you estimate a bivariate regression model with a total sample size of 50 and obtain a standard error for $\beta_1$ of 2.50. What is the smallest regression coefficient ($\beta_1$) that you could have and still be able to reject the null hypothesis (of no relationship between $X_1$ and $Y$) at the 0.05 level with a one-tailed test?

**The smallest regression that allows for rejection at is 4.19.**

### Question 2

I regress $Y$ on $X_1$ (i.e., $Y = \beta_0 + \beta_1 X_1$) and find that $\beta_1 = 4.20$, $SE_{\beta_1} = 1.90$, $t = 2.21$, $p = 0.01$, and a 95% confidence interval is \[0.48, 7.92\]. What is my best estimate of the effect of a two-unit change in $X_1$ on the mean of $Y$ in the population?

**The best estimate is 8.40.**

### Question 3

I conduct an OLS regression with a sample size of 90 and 5 independent variables. To determine a p-value for each coefficient, I would examine a t-distribution with how many degrees of freedom?

**I would examine a t-distribution of 84 degrees of freedom.**

### Question 4

Suppose I regress $Y$ on $X$ and compute the mean response for $Y$ at some specified value of $X$. When determining the confidence interval around this mean response, which of the following will **NOT** have any effect on the width (or size) of that confidence interval?

**D**

A.  Total sample size,

B.  Mean-squared error,

C.  The specified value of $X$,

D.  **None of the above: all of these (above) will affect the confidence interval,**

E.  There is not enough information to answer this question.

### Question 5

I regress $Y$ on $X$ and find that $\beta_1$ has a two-tailed p-value of 0.04. Which of the following statements is most accurate?

**D**

A.  The lower and upper bound of a 95% confidence interval around $\beta_1$ will have the same sign,

B.  The absolute value for the t-statistic for $\beta_1$ will be greater than 1.96,

C.  A 90% confidence interval around $\beta_1$ will not contain zero,

D.  **All of the above**

E.  None of the above and/or there is not enough information.

### Question 6

I regress $Y$ on three independent variables -- $X_1$, $X_2$, and $X_3$ -- and I find the following 95% confidence intervals -- $\beta_1$: \[0.12, 1.45\], $\beta_2$: \[-0.01, 0.15\], and $\beta_3$: \[-0.64, -0.01\]. Which of the following statements is most accurate?

**E**

A.  $\beta_2$ and $\beta_3$ will have negative coefficients, and $\beta_1$ is statistically significant at the 0.05 level (two-tailed),

B.  $\beta_1$ will have a positive coefficient, and $\beta_1$ is the only statistically significant coefficient (of the three coefficients) at the 0.05 level (two-tailed),

C.  $\beta_1$ will have a positive coefficient, $\beta_2$ and $\beta_3$ will have negative coefficients, and only $\beta_1$ and $\beta_3$ are statistically significant at the 0.05 level (two-tailed),

D.  All of the above,

E.  **None of the above and/or there is not enough information.**

### Question 7

Suppose I estimate a regression with two independent variables and obtain a $R^2$ of 0.40 where the Residual Sum of Squares is equal to 5,150. What does the Total Sum of Squares equal in this regression model?

**The TSS is 8,583.33.**

### Question 8

I regress $Y$ on $X_1$ and $X_2$ (i.e., $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2$). I find that $\beta_1 = -0.87$ with a two-tailed p value of 0.001 and $\beta_2 = 3.52$ with a two-tailed p-value of 0.04. Which of the following statements is most accurate?

**A**

A.  $X_2$ **has a larger substantive effect on** $Y$ **than** $X_1$**, but the effect of** $X_1$ **is more statistically significant**

B.  $X_1$ has a larger substantive effect on $Y$ than $X_2$ and $X_1$ is more statistically significant than $X_2$,

C.  $X_1$ has a larger substantive effect on $Y$ than $X_2$, but the effect of $X_2$ is more statistically significant,

D.  None of the above and/or there is not enough information.

### Question 9

I regress $Y$ on $X_1$ (i.e., $Y = \beta_0 + \beta_1 X_1$) and obtain a $R^2$ of 0.45. Then, I regress $Y$ on both $X_1$ and $X_2$ (i.e., $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2$) and obtain a $R^2$ of 0.53. Which of the following statements must be true?

**C**

A.  The correlation between $X_1$ and $Y$ is stronger than the correlation between $X_2$ and $Y$,

B.  The coefficient on $\beta_1$ is statistically significant in both models,

C.  **The model with** $X_1$ **and** $X_2$ **explains more of the variation in** $Y$ **than a model with just** $X_1$**,**

D.  All of the above: all of these statements are true.

### Question 10

Based on the following regression equation, might the negative coefficient on the South variable be the result of southern states having a lower proportion of high school graduates than non-southern states? In no more than one sentence, explain your answer.

$$
Turnout = 34 + 0.5(Percent\ High\ School\ Graduates) – 5.9(South)
$$

**Yes, since the regression already takes into account the influence of the percentage of high school graduates, it is possible that the negative coefficient on the south variable is caused by the lower proportion of high school graduates in southern states. This would isolate the distinct contribution of being a southern state to lower turnout.**

## Applied Questions

All datasets referenced below are posted on the ELMs course site.

### Question 1

*Points: 15*

Use the `world` dataset to answer the following questions. The relevant variables and their coding information are as follows: `literacy` indicates a country's literacy rate; `dem_score14` represents a country's level of democratization (higher values indicate greater democratization); `spendeduc` reflects the amount of public expenditures on education as a percentage of GDP; `gdp_10_thou` represents GDP per capita (one unit is \$10,000 USD); `educ_quality` indicates the average quality rating of a country's educational system; and `ungr9095` represents the percent average annual population growth.

#### Part A

*Points: 4*

Evaluate the bivariate relationship (using R) between a country's support for public education and its literacy rate. Report the results in a professionally formatted table. Do countries that spend more money on education appear to have a higher literacy rate?

```{r}

world <- read_csv("world.csv")

summary(lm(literacy ~ spendeduc, data = world))

ggplot(world, aes(x = spendeduc, y = literacy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Public Education Spending and Literacy Rate",
       x = "Public Education Spending (% of GDP)",
       y = "Literacy Rate (%)")

model <- lm(literacy ~ spendeduc, data = world)
stargazer(model, type = "text", title = "Bivariate Relationship between Public Education Spending and Literacy Rate",
          dep.var.labels = "Literacy Rate", covariate.labels = "Public Education Spending (% of GDP)")
```

**The data would suggest that countries that spend more on education have higher literacy rates.**

#### Part B

*Points: 5*

Estimate a second regression model (using R) that includes both a country's public expenditures on education and its level of democratization as independent variables. When controlling for democratization, does education expenditures exhibit a significant impact on literacy? Report the results in a professionally formatted table. Be sure to discuss the extent to which the results change, and if so, why they changed.

```{r}
world_data_clean <- na.omit(world[, c("spendeduc", "literacy", "dem_score14")])
summary(world_data_clean)

model2 <- lm(literacy ~ spendeduc + dem_score14, data = world_data_clean)
summary(model2)

stargazer(model2, type = "text", title = "Regression: Literacy Rate vs. Education Spending and Democratization",
          dep.var.labels = "Literacy Rate",
          covariate.labels = c("Public Education Spending", "Democratization Level"))
```

**Education expenditures exhitbit a signficicant impcat on literacy. Including democratization changes the results because it captures a explanatory variable for literacy rates. Democratization likely mediates or confounds the relationship between education spending and literacy.**

#### Part C

*Points: 6*

Now estimate a third model (using R) that includes all of the variables listed above. Report the regression results in a professionally formatted table and interpret each coefficient. When interpreting the impact of democratization using this regression model, present/utilize a visual representation (using R) of the expected change in literacy as a function of democratization.

```{r}
variables <- c("literacy", "spendeduc", "dem_score14", "gdp_10_thou", "educ_quality", "ungr9095")

world_data_clean <- na.omit(world[, variables])

summary(world_data_clean)

model3 <- lm(literacy ~ spendeduc + dem_score14 + gdp_10_thou + educ_quality + ungr9095, data = world_data_clean)
summary(model3)

stargazer(model3, type = "text",
          title = "Regression Results: Literacy Rate vs. All Independent Variables",
          dep.var.labels = "Literacy Rate",
          covariate.labels = c("Public Education Expenditures (%)", "Level of Democratization",
                               "GDP per Capita (in $10,000s)", "Education Quality", 
                               "Population Growth Rate (1990-95)"))
visual_data <- data.frame(
  dem_score14 = seq(min(world_data_clean$dem_score14), max(world_data_clean$dem_score14), length.out = 100),
  spendeduc = mean(world_data_clean$spendeduc),
  gdp_10_thou = mean(world_data_clean$gdp_10_thou),
  educ_quality = mean(world_data_clean$educ_quality),
  ungr9095 = mean(world_data_clean$ungr9095)
)
visual_data$predicted_literacy <- predict(model3, newdata = visual_data)

ggplot(visual_data, aes(x = dem_score14, y = predicted_literacy)) +
  geom_line(color = "blue", size = 1.2) +
  labs(title = "Predicted Literacy as a Function of Democratization",
       x = "Democratization Score",
       y = "Predicted Literacy Rate") +
  theme_minimal()
```

### Question 2

*Points: 15*

Using the `states` dataset, use R to regress the variable `obama08` (the percentage of a state's vote that President Obama received in the 2008 U.S. presidential election) on the following independent variables: `cig_tax` represents the amount of a state's cigarette tax (in dollars); `college` is the percentage of a state's population that graduated college (0 to 100); `union07` is the percentage of a state's workers that are union members (0 to 100); and `south` is a dichotomous variable coded as `1` if the state is located in the South (`0` otherwise). Report the results in a professionally formatted table. Answer the following questions.

#### Part A

*Points: 6*

Interpret the substantive results from this regression model. Be sure to discuss the direction, magnitude, and statistical significance of each slope coefficient. Does the `union07` variable exhibit a substantively significant impact on voting behavior in the 2008 election?

```{r}
states <- read_csv("states.csv")

model <- lm(obama08 ~ cig_tax + college + union07 + south, data = states)
summary(model)

stargazer(model, type = "text",
          title = "Regression Results: Obama 2008 Vote Share vs. State-Level Factors",
          dep.var.labels = "Obama 2008 Vote Share (%)",
          covariate.labels = c("Cigarette Tax ($)", "College Graduates (%)",
                               "Union Membership (%)", "South (1 = Yes)"),
          out = "results.txt")
```

**The coefficient for cigarette tax is positive, indicating that states with higher cigarette taxes tend to have a higher vote share for Obama. For each additional dollar in cigarette tax, Obama's vote share increases by approximately 1.5 percentage points. The p-value is less than 0.05 making the relationship statistically significant.**

**The coefficient for college graduates is positive, showing that a higher percentage of college graduates in a state is associated with a higher vote share for Obama. For each 1% increase in the percentage of college graduates, Obama's vote share increases by approximately 0.8 percentage points. The p-value is less than 0.01 indicating strong statistical significance.**

**The coefficient for union membership is positive, suggesting that states with a higher percentage of unionized workers tend to have a higher vote share for Obama. For each 1% increase in union membership, Obama's vote share increases by approximately 0.4 percentage points. The p-value is greater than 0.05 meaning the relationship is not statistically significant.**

**The coefficient for south is negative, indicating that states located in the South had a lower vote share for Obama compared to non-southern states. Being in the South is associated with a 10-percentage-point decrease in Obama's vote share. The p-value is less than 0.01 making the relationship statistically significant.**

**The union membership variable does not exhibit a substantively significant impact on voting behavior.**

#### Part B

*Points: 2*

Interpret the intercept coefficient. What does this represent and is it meaningful to interpret this coefficient on its own? Why or why not?

**The intercept in this regression model is not meaningful to interpret on it's own because it i unrealistic for the many variables to have a 0% condition.**

#### Part C

*Points: 2*

What is the expected mean percentage of Obama's vote total in a Southern state with a cigarette tax of \$2.00, a state population where 30% of people are college graduates, and where 15% of workers in a state are union members?

```{r}
50 + 3 + 24 + 6 - 10 
```

**The expected mean percentage is 73% in a southern state.**

#### Part D

*Points: 3*

What is a 90% confidence interval around the coefficient for the `south` variable.

```{r}
coef_south <- coef(summary(model))["south", "Estimate"]
se_south <- coef(summary(model))["south", "Std. Error"]

df <- model$df.residual
t_crit <- qt(0.95, df)

lower_bound <- coef_south - t_crit * se_south
upper_bound <- coef_south + t_crit * se_south

cat("90% Confidence Interval for the 'south' variable:", lower_bound, "to", upper_bound, "\n")
```

**The 90% confidence interval \[-2.95, 5.73\].**

#### Part E

*Points: 3*

What is the null hypothesis that the F-test in this regression output is testing?

**The null hypothesis is testing if all the independent variables regression coefficients are all equal to zero at the same time.**
