---
title: "Mid-term Exam #2"
subtitle: "Due date: 6 November"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---

**Kennedy Lighty**

```{r}
library(tidyverse)
library(peacesciencer)
library(ggdist)
library(MetBrewer)
library(scales)
library(janitor)
library(wbstats)
library(countrycode)
library(modelsummary)

```

Please read all of the questions carefully and follow all instructions. Each question has an allotted point value. Be as thorough and specific as possible; extra calculations and incorrect information, even in the presence of correct information, will result in point deductions. Be sure to show all formulas and all necessary work to answer the questions. You may upload your completed exam to the Elms course site (attach to Midterm Exam #2).

::: callout-note
While this is an open-note exam, you are not to receive assistance from anyone else (as usual, the Honor Code applies).
:::

**Total points: 50 points**

## Conceptual questions

::: callout-note
Please include all work (and computations) necessary to answer the questions.
:::

**Total points: 26**

### Question 1

*2 points*

::: callout-note
2/2
:::

Suppose I am interested in determining if freshman undergraduates at the University of Maryland spend more hours studying in the average week than sophomore undergraduates. I conduct a study in which I take a simple random sample (SRS) of 1200 freshman students and 1200 sophomore students. I find that the freshman students in my sample study for, on average, 412 minutes per week and the sophomore students in my sample study for, on average, 335 minutes per week. The standard error of the difference is 30. What is a 90% confidence interval for the difference between freshman and sophomore students?

**The 90% confidence interval for students is 27.65, 126.35**

### Question 2

*2 points*

::: callout-note
1/2
:::

Based on the results of my study described in question 1, can you reject the null hypothesis of no statistically meaningful difference in the study habits of sophomore and freshman students? Why or why not?

::: callout-note
This suggests that we can reject the null hypothesis.
:::

**We can not reject the null hypothesis because the interval does not include 0 it suggest that there is a significant statistical difference between the two.**

### Question 3

*2 points*

::: callout-note
1/2
:::

If I am testing a null hypothesis that X has no effect on Y in the population (and thus my alternative hypothesis is that X does have an effect), why might I prefer to commit a Type-II error over a Type-I error (and, of course, this holds aside my first preference of making no error at all)? Answer in no more than two sentences.

::: callout-note
These are definitions, not an argument for your preference.
:::

**I would prefer a type II error because it tells us that there is simply no effect. A type I error is worse because it indicates a potential false positive.**

### Question 4

*1 point*

::: callout-note
1/1
:::

When conducting a difference-of-means test, which of the following samples would yield a sampling distribution with the lowest variability?

A.  **A sample of 900 with a standard deviation of 15**

B.  A sample of 25 with a standard deviation of 10

C.  A sample of 625 with a standard deviation of 20

D.  A sample of 100 with a standard deviation of 6

### Question 5

*1 point*

::: callout-note
1/1
:::

Which of the following probabilities is not independent?

**C. The probability that I draw an ace and then a king in a row from a deck of cards (when drawing a two-card hand).**

A.  The probability that the roulette wheel will end up on 23 two times in a row.
B.  The probability that three successive coin tosses will each turn up heads.
C.  **The probability that I draw an ace and then a king in a row from a deck of cards (when drawing a two-card hand).**
D.  The probability that I will get a 6 on three subsequent die rolls.
E.  None of the above – all are independent probabilities.

### Question 6

*1 point*

::: callout-note
1/1
:::

I conduct a two-tailed difference-of-means test and obtain a t-statistic of 2.10. Is my result statistically significant (with 95% confidence)?

**D. There is not enough information to answer this question.**

A.  Yes, at the 0.05 level (p\<0.05)
B.  Yes, but only at the 0.10 level (p\<0.10)
C.  Not at either the 0.05 or the 0.10 level
D.  **There is not enough information to answer this question.**

### Question 7

*1 point*

::: callout-note
1/1
:::

I take a sample of 1800 adults and find that 360 of them watched last Monday’s NFL game. What probability represents the complement to the sample proportion of adults who watched the NFL game?

**The percentage of adults in the sample who watched the NFL game is 0.8. The likelihood that an adult in the sample did not watch the game is represented by this.**

### Question 8

*1 point*

::: callout-note
0/1, E.
:::

Which of the following makes it more likely that a given sample statistic will be statistically different from zero (and thus allow you to reject the null hypothesis, all else equal)?

**A. More observations**

A.  **More observations**
B.  Greater variance in the sample
C.  Using a two-tailed instead of a one-tailed significance test
D.  A larger confidence interval around the test statistic
E.  Both (a) and (b) (but not (c) or (d))
F.  All of the above increase the chances of statistical significance
G.  None of the above

### Question 9

*1 point*

::: callout-note
1/1
:::

The p-value for a two-tailed test of the null hypothesis $H_0: \mu=40$ is 0.06. Which of the following statements is accurate?

**A. A 95% confidence interval around the sample mean includes the value 40**

A.  **A 95% confidence interval around the sample mean includes the value 40**
B.  A 90% confidence interval around the sample mean includes the value 40
C.  A 99% confidence interval around the sample mean does not include the value 40
D.  None of the above statements are correct
E.  All of the above statements are correct

### Question 10

*4 points total*

#### Part A

*2 points*

::: callout-note
0/2
:::

If you roll a fair (six-sided) die twice, what is the probability that both rolls will be odd or greater than four?

**The probability that both rolls will be odd or greater than 4/9th.**

::: callout-note
The probability that both rolls will be odd:

$$
Pr(Odd, Odd) = \frac{3}{6}*\frac{3}{6} = 0.25
$$

The probability that both rolls will be greater than four:

$$
Pr(>4, >4) = \frac{2}{6} * \frac{2}{6} = 0.111
$$

The probability that both rolls will be both odd and greater than four:

$$
Pr(5, 5) = \frac{1}{6} * \frac{1}{6} = 0.028
$$

Therefore:

$$
Pr(Odd, Odd\ OR >4, >4) = Pr(Odd, Odd) + Pr(>4, >4) - Pr(5, 5) = 0.25 + 0.111 - 0.028 = 0.333
$$
:::

#### Part B

*2 points*

::: callout-note
2/2
:::

The following is a distribution of U.S. college students classified by their age and full- vs. part-time status. Based on these data, what is the probability that a student is in the 25-29 age group, given that (i.e., conditional on knowledge that) the student is full time?

```{r}
tibble::tibble(
  age = c("15 - 19", "20 - 24", "25 - 29", "30+"),
  full_time = c(155, 255, 75, 35),
  part_time = c(20, 55, 80, 95)
) |> 
  knitr::kable()
```

**The probability is 14.4.**

### Question 11

*4 points total*

#### Part A

*2 points*

::: callout-note
2/2
:::

Using a SRS of 1211 people, I estimate that the proportion of people living in the South that favor teaching sexual education in public schools is 0.874 and the proportion of people outside of the south that favor teaching sexual education in public schools is 0.906. And, the standard error of the difference (in citizen views about teaching sexual education in public schools) between people living in the south and those not living in the south is 0.015. Give an interval estimate for the difference in the proportion of people favoring sex education in public schools between people who do, and do not, live in the south.

**The interval estimate for the difference in proportion is (-0.0614, -0.0026).**

#### Part B

*2 points*

::: callout-note
2/2
:::

Do the data (i.e., estimates above) show support for my hypothesis? How do you know?

**The data supports the hypothesis because of the statistically significant difference shown by the confidence interval below zero.**

### Question 12

*6 points total*

I am interested in estimating the average number of texts that University of Maryland undergraduate students send in a day. My hypothesis is that the average number of texts students send is greater than 100. I take a SRS of 1600 students and find that the mean number of texts they send is 105, and with a standard deviation of 120.

#### Part A

*2 points*

::: callout-note
2/2
:::

What is a 95% confidence interval around the sample statistic?

**The 95% interval is (99.12, 110.88)**

#### Part B

*2 points*

::: callout-note
0/2
:::

When testing the null hypothesis, what is the test statistic associated with the sample statistic?

::: callout-note
This is a means test, so you need the t-statistic.
:::

**The Z-score is 1.67**

#### Part C

*2 points*

::: callout-note
2/2, error carried forward.
:::

If using a one-tailed test of the null hypothesis and you are willing to accept a Type-I error rate of 0.05, do the data support my hypothesis? Why or why not?

**The data supports the hypothesis because 1.67 is greater than the critical value 1.64.**

## Applied questions

::: callout-note
All datasets referenced below are available through the `poliscidata` package.
:::

**Total points: 24**

### Question 1

```{r}
gss <- poliscidata::gss |>

mutate(voted08 = as.numeric(voted08),

conarmy = as.numeric(conarmy))
```

*8 points total*

I hypothesize that, among only those that were eligible to vote, people with greater confidence in the U.S. military were more likely to turnout to vote in the 2008 presidential election. Use data from the General Social Survey (i.e., the `gss` dataset) to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `conarmy` (1 = a “great deal” of confidence; 2 = “only some” confidence; 3 = “hardly any”); and `vote08` (1 = did not vote; 2 = voted). Answer the following questions.

#### Part A

*3 points*

::: callout-note
1/3
:::

Complete a cross-tab and interpret the results. Do the data support my hypothesis? Be sure to explain the nature of the relationship (or lack thereof, if relevant).

```{r}
gss_filtered <- gss %>% filter(!is.na(voted08) & !is.na(conarmy))

datasummary_crosstab(
  formula = voted08 ~ conarmy,
  data = gss_filtered,
  output = "markdown"
)
```

::: callout-note
You need to calculate the proportions within your IV groups (down the columns).
:::

**The data supports the hypothesis because there appears to be a positive correlation between rising military confidence and increased voter turnout.**

#### Part B

*3 points*

::: callout-note
3/3
:::

Compute (by hand) the chi-square statistic to test the null hypothesis of no relationship between these two variables. Be sure to show your work.

```{r}
gss_filtered <- gss %>% filter(!is.na(voted08) & !is.na(conarmy))
observed <- table(gss_filtered$voted08, gss_filtered$conarmy)
observed_df <- as.data.frame(as.table(observed))
colnames(observed_df) <- c("Voted08", "Conarmy", "Observed")
row_totals <- margin.table(observed, 1)
col_totals <- margin.table(observed, 2)
grand_total <- sum(observed)

observed_df <- observed_df %>%
  mutate(Expected = (row_totals[Voted08] * col_totals[Conarmy]) / grand_total,
         ChiSquare = (Observed - Expected)^2 / Expected)
chi_square_stat <- sum(observed_df$ChiSquare)
chi_square_stat
```

#### Part C

*2 points*

::: callout-note
2/2
:::

Using the chi-square statistic that you computed in question 1(b), can you reject the null hypothesis of no relationship between these two variables with 95% confidence? Why, or why not?

**Based on the Chi-square stat of 6.72 we can reject the null hypothesis because a state of 6.72 shows a statistically significant relationship between confidence.**

### Question 2

nes \<- poliscidata::nes \|\> dplyr::mutate(envir_drill = factor(envir_drill, levels = c("1. Favor", "3. Neither favor nor oppose", "2. Oppose")), envir_drill = as.numeric(envir_drill), pid_x = as.numeric(pid_x), relig_pray = as.numeric(relig_pray), voted2008 = as.numeric(voted2008))

```{r}

nes <- poliscidata::nes |> 
  dplyr::mutate(envir_drill = factor(envir_drill, levels = c("1. Favor",
                                                             "3. Neither favor nor oppose",
                                                             "2. Oppose")),
                envir_drill = as.numeric(envir_drill),
                pid_x = as.numeric(pid_x),
                relig_pray = as.numeric(relig_pray),
                voted2008 = as.numeric(voted2008))
```

*8 points total*

I hypothesize that citizens who do not support increased U.S. offshore drilling are more conservative than those who do not. Use data from the `nes` dataset to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `envir_drill` (1 = “favor”; 2 = “neither favor nor oppose”; 3 = “oppose”); and `pid_x` (higher values represent less liberalism, or more conservatism). Answer the following questions.

#### Part A

*2 points*

::: callout-note
2/2
:::

Using these data, what is the point estimate for the mean conservatism/liberalism score among those that oppose drilling? What is the point estimate for the mean conservatism/liberalism score among those that are in favor of drilling?

```{r}
nes_filtered <- nes %>% filter(!is.na(envir_drill) & !is.na(pid_x))

point_estimates <- nes_filtered %>%
  group_by(envir_drill) %>%
  summarize(mean_conservatism = mean(pid_x, na.rm = TRUE))

point_estimates
```

**For those who are opposed to off shore drilling the point estimate is 2.63 and those who are for the point estimate is 4.17.**

#### Part B

*6 points*

::: callout-note
0/6
:::

Evaluate the null hypothesis that there is no difference in the mean conservatism/liberalism score among those that approve vs. disapprove of drilling. Do the data support my hypothesis? Why or why not? Be sure to show all work necessary to answer the question by hand (i.e., you may only use R to the extent that is absolutely necessary to complete the question; otherwise, you must show how you would answer the question by hand).

### Question 3

*8 points total*

I hypothesize that people who express that religion is important to them were more likely to turnout to vote in the 2008 presidential election. Use data from the `nes` dataset to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `relig_pray`, which describes how often the respondent prays (1 = several times a day; 2 = once a day; 3 = a few times a week; 4 = once a week or less; 5 = never); and `voted2008` (1 = did not vote; 2= voted). Answer the following questions.

#### Part A

*2 points*

::: callout-note
2/2
:::

Using these data, what is the point estimate for the proportion of respondents that voted (i.e., turnout rate), among citizens expressing that religion is not important? What is the point estimate for the proportion of respondents that voted, among citizens expressing that religion matters a great deal?

```{r}
nes %>%
  filter(!is.na(relig_pray) & !is.na(voted2008)) %>%
  mutate(
    relig_importance = case_when(
      relig_pray %in% c(1, 2) ~ "Matters a Great Deal",
      relig_pray == 5 ~ "Not Important",
      TRUE ~ "Other"
    )
  ) %>%
  filter(relig_importance %in% c("Matters a Great Deal", "Not Important")) %>%
  group_by(relig_importance) %>%
  summarise(turnout_rate = mean(voted2008 == 2, na.rm = TRUE))
```

**The point estimate for those who do not believe religion matters is 0.72. the point estimate for those who believe religion matters is 0.81.**

#### Part B

*6 points*

::: callout-note
0/6
:::

Evaluate the null hypothesis that there is no difference in the proportion of voters (i.e., turnout rate) among citizens expressing that religion is not important vs. those reporting that religion matters a great deal. Do the data support my hypothesis? Why or why not? Be sure to show all work necessary to answer the question by hand (i.e., you may only use R to the extent that is absolutely necessary to complete the question; otherwise, you must show how you would answer the question by hand).
