---
title: "Problem Set 5"
subtitle: "Due date: 23 October"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
---

**Kennedy Lighty**

```{r}
library(tidyverse)
library(janitor)
library(poliscidata)
library(ggdist)
library(MetBrewer)
library(modelsummary)
library(DescTools)
library(scales)
```

Please upload your completed assignment to the ELMs course site (under the assignments menu) and to your class Github repository. You need to upload both your Quarto document (with R code, as needed) and its rendered output (either as a PDF or HTML file).

**Total points: 25**

## Question 1

*Total points: 6*

Use the data in the table below to answer the following questions.

```{r}
#| echo: false

tibble::tibble(
  `Age group` = c("18-24", "25-30", "31 and up", "TOTAL"),
  `Non-voters` = c(70, 40, 220, 330),
  Voters = c(50, 50, 570, 670),
  Total = `Non-voters` + Voters
) |> 
  knitr::kable(caption = "Voting by Age in 2000")
```

### Part A

*Points: 2*

::: callout-note
2/2
:::

What is the probability of being 25-30 or a non-voter?

```{r}
90/1000 # 25-30
330/1000 #non voter
40/1000 #non voter and 25-30

0.09 + 0.33 - 0.04
```

**38%**

### Part B

*Points: 4*

::: callout-note
1/4
:::

Assuming a normal distribution, report the 95% confidence intervals for the percentage of 18-to-24-year-olds who did not vote, and then the percentage of 25-to-30-year-olds who did not vote.

```{r}
70/120

0.5833333 * (1 - 0.5833333)

0.2430556/120
sqrt(0.002025463)

0.5833333 - 1.96 * 0.04500514


40/90

0.4444444 - 1.96 * sqrt((0.4444444*(1 - 0.4444444))/90)
0.4444444 + 1.96 * sqrt((0.4444444*(1 - 0.4444444))/90)
```

**Age Range -18 - 24 year: 49.5% and 67.2%**

**Age Range- 25-30: 34.2% and 54.7%**

::: callout-tip
The probability that an individual is 18-to-24 and a non-voter is not conditional. Therefore, you look at the proportion of the whole population who meet those criteria.

$$
\hat{p} = Pr(18\ to\ 24 \& NV) = \frac{70}{1000} = 0.07 
$$

And:

$$
n = 1000
$$

Therefore:

$$
SE(\hat{p}) = \sqrt{\frac{0.07 * 0.93}{1000}} = 0.008068457
$$

The 95% confidence interval can be found:

$$
\hat{p} \pm 1.96*SE(\hat{p})
$$

Therefore, the 95% confidence interval for the percentage of 18-to-24-year-olds who did not vote is `r percent(0.07 - 1.96 * 0.008068457, accuracy = 0.01)` to `r percent(0.07 + 1.96 * 0.008068457, accuracy = 0.01)`.

Using the same method, the 95% confidence intervals for the percentage of 25-to-30-year-olds who did not vote is `r percent(0.04 - 1.96*0.006196773, accuracy = 0.01)` to `r percent(0.04 + 1.96*0.006196773, accuracy = 0.01)`.
:::

## Question 2

*Total points:* *7*

Assume that the standard deviation for the population distribution of a state in which you want to conduct a poll is 200.

### Part A

*Points: 3*

::: callout-note
3/3
:::

Calculate the spread of the sampling distribution for each of the following sample sizes: 1, 4, 25, 100, 250, 1000, 5,000, and 10,000.

```{r}
200/sqrt(1)

200/sqrt(4)

200/sqrt(25)

200/sqrt(100)

200/sqrt(250)

200/sqrt(1000)

200/sqrt(5000)

200/sqrt(10000)
```

**The sample sizes are 200, 100, 40, 20, 12.64, 6.32, 2.82, and 2**

### Part B

*Points: 1*

::: callout-note
1/1
:::

Describe specifically how the variability of the sampling distribution changes as the sample size varies. Considering the expense of running a poll, which sample size do you think is most optimal if conducting the poll?

**As sample size increases, the distribution's spread gets closer together. A sample size of 1,000–1,500 is frequently ideal for conducting a poll. This size strikes a compromise between cost and precision, offering a modest estimate that is relatively correct.**

### Part C

*Points: 3*

::: callout-note
3/3
:::

Display your results graphically (using R) with the sample size on the x-axis and the standard error (of the sampling distribution) on the y-axis.

```{r}
sample_size <- c(1, 4, 25, 100, 250, 1000, 5000, 10000)

standard_error <- c(200, 100, 40, 20, 12.65, 6.32, 2.83, 2)

data <- data.frame(
  sample_size = c(1, 4, 25, 100, 250, 1000, 5000, 10000),
  standard_error = c(200.0, 100.0, 40.0, 20.0, 12.65, 6.32, 2.83, 2.0)
)

ggplot(data, aes(x = sample_size, y = standard_error)) + 
   geom_line(color = "pink") + 
  geom_point(color = "red") +
  theme_minimal() + 
  labs( title = "Comparing Sample Size and Standard Error",
       x = "Number of observations in the sample",
       y = "Standard error")
```

## Question 3

*Points: 4*

::: callout-note
4/4
:::

Suppose you conduct a survey (to generate a sample mean of interest) and find that it has a margin of error of 4.5 with a sample size of 900 using a 95% confidence interval. What would the margin of error be for a 90% confidence interval?

```{r}
4.5/1.96

1.645 * 2.29
```

**3.77**

## Question 4

*Points: 4*

::: callout-note
4/4
:::

Assume that, in State A, the mean income in the population is \$20,000 with a standard deviation of \$2,000. If you took an SRS of 900 individuals from that population, what is the probability that you would get a sample mean income of \$20,200 or greater? What would be the probability if the sample size was only 25?

::: callout-note
Assume a normal distribution for both questions.
:::

```{r}
2000/sqrt(900)

(20200-20000)/66.66667

2000/sqrt(25)

(20200-20000)/400
```

**900: 0.00135 \| 25: 0.3085**

## Question 5

*Points: 4*

::: callout-note
4/4
:::

Assume that a coin is fair. If I flip a coin 500 times, what is a 95% confidence interval for the range of the count of heads that I will get? What if I flip the coin 5,000 times? What about 50,000 times?

```{r}
sqrt(500 * .5 * (1 - .5)) #standard deviation 1

500 * .5 #mean 1

5000 * .5

50000 * .5

sqrt(5000 * .5 * (1 - .5))

sqrt(50000 * .5 * (1 - .5))

250 + 1.96 * 11.18

250 - 1.96 * 11.18

2500 + 1.96 * 35.36

2500 - 1.96 * 35.36

25000 + 1.96 * 111.80

25000 - 1.96 * 111.80
```

**500: 271.91, 228.09 \| 5000: 2569.31, 2430.70 \| 50,000: 25219.12, 24780.87**
