---
title: "Problem Set 6"
subtitle: "Due date: 23 October"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
  warning: false
---

**Kennedy Lighty**

```{r}
library(tidyverse)
library(skimr)
library(wbstats)
library(countrycode)
library(janitor)
library(broom)
library(infer)
library(scales)

```

Please upload your completed assignment to the ELMs course site (under the assignments menu) and to your class Github repository. You need to upload both your Quarto document (with R code, as needed) and its rendered output (either as a PDF or HTML file).

**Total points: 35**

## Question 1

*Points: 10*

I hypothesize that the mean mathematics SAT score for University of Maryland students is different to 600. I take an SRS of 625 students and find that the mean mathematics score among those students is 610. The standard deviation of my sample is 75.

### Part A

::: callout-note
2/2
:::

Based on my sample, what is a point estimate for the mean mathematics SAT score for University of Maryland students?

**Based on the sample the estimate score is 610.**

### Part B

::: callout-note
2/2
:::

What is a 95% confidence interval around that estimate?

**Based on the sample the 95% interval is 604.12% and 615.88%**

### Part C

::: callout-note
2/2
:::

What is the null hypothesis?

**The null hypothesis is 600.**

### Part D

::: callout-note
0/2
:::

What is the two-tailed p-value and what does it mean?

**The two tailed value is 0.86. This means that the probability of observing a sample mean is more than likely between 610 and 600.**

::: callout-note
p-value: 0.0009

It means that there is a vanishingly small chance (\<0.001%) that I would observe a value that was either 10 points higher or lower than a mean mathematics SAT score for University of Maryland students of 600.

Please show your working (for the p-value). I can't give you points if I don't know what you did.
:::

### Part E

::: callout-note
0.5/2
:::

Do the data support my hypothesis? Why, or why not?

**Yes, the data does support my hypothesis. There is considerable statistical support for the hypothesis that University of Maryland students actual mean SAT score in mathematics is lower than 600.**

::: callout-note
Correct, but I don't know how you got there. A p-value of 0.86 would not show support for the hypothesis.
:::

## Question 2

*Points: 10*

I take an SRS of 900 citizens in a public opinion survey and find that the mean feeling thermometer rating (on a 0 to 100 scale) of President Biden is 51.5 with a standard deviation of 25. Now, suppose I hypothesize that feelings toward Biden should be greater than 50.

### Part A

::: callout-note
0/2
:::

What is the null hypothesis?

**The null hypothesis is less than 50.**

::: callout-note
Null hypothesis is 50. It is always a single number (not a range).
:::

### Part B

::: callout-note
2/4
:::

What is a 95% confidence interval around the sample statistic? Does this interval indicate support for a two-tailed test of my hypothesis (`p < 0.05` as the threshold)? Why, or why not?

**The 95% confidence interval is 49.87% and 53.13%. This does not indicate strong support of the hypothesis because it is less than 50.**

::: callout-note
... because it includes the null hypothesis of 50.
:::

### Part C

::: callout-note
3.5/4
:::

What is a 90% confidence interval around the sample statistic? Does this interval indicate support for a one-tailed test of my hypothesis (`p < 0.05` as the threshold)? Why, or why not?

**The 90% confidence interval is 50.13% and 52.87%. This does indicate a strong support of the hypothesis because it is above 50.**

::: callout-note
... because it does not include the null hypothesis of 50.
:::

## Question 3

*Points: 5*

::: callout-note
5/5
:::

Suppose I conduct a difference of means test and obtain a t-statistic of 2.50. What does this indicate about the statistical significance of the sample mean compared to the null hypothesis?

**This indicates that the sample mean is significantly different from the null hypothesis.**

## Question 4

*Points: 5*

I hypothesize that fewer than 40% of registered voters approve of President Biden. I conduct an SRS of 720 registered voters and find that 37.9% of them approve of President Biden.

### Part A

::: callout-note
1/1
:::

Based on my sample, what is a point estimate for the proportion of the population that approves of President Biden?

**The point estimate is 37.9%.**

### Part B

::: callout-note
1/1
:::

What is a 90% confidence interval around that estimate?

**The 90% confidence interval is 34.93 and 40.87**

### Part C

::: callout-note
0/1
:::

What is the null hypothesis?

**The null hypothesis is more than 40%**

::: callout-note
Is equal to 40%.
:::

### Part D

::: callout-note
0.5/1
:::

What is the one-tailed p-value and what does it mean?

**The one-tailed p value is 0.12. This means the data does not support the null hypothesis.**

::: callout-note
This means that there is a 12% chance that we would observe this sample proportion or smaller if the null hypothesis were true.
:::

### Part E

::: callout-note
1/1
:::

Do the data support my hypothesis? Why, or why not?

**The data does not support the hypothesis because the p-value is greater than the 0.05. The data does not show evidence that is statistically significant to back up the hypothesis.**

## Question 5

*Points: 5*

::: callout-note
0/5
:::

```{r}

library(wbstats)
library(tidyverse)

q_5_df <- wb_data(c("SP.DYN.TFRT.IN", "SE.ADT.1524.LT.FE.ZS"),
                  start_date = 2022,
                  end_date = 2022) |> 
  rename(lit_rate = `SE.ADT.1524.LT.FE.ZS`,
         fertility = `SP.DYN.TFRT.IN`) |> 
  mutate(lit_rate = cut_interval(lit_rate, 2, labels = c("Lower half", "Upper half"))) |> 
  drop_na(fertility, lit_rate)
```

I hypothesize that countries with higher rates of literacy among young women have lower birth rates, on average, compared to countries that have lower rates of young female literacy. Use data collected above from the World Bank to test my hypothesis. State the null hypothesis and interpret what the results tell us about the null and alternative hypotheses. Show all relevant work in a professional format.
